Sketching for Density Estimation

### 2018-12-09 density.tex
small changes + bugfix

\section{Appendix 1}
For completeness we recap a result from \cite{barany2008}.
\begin{theorem}[Simplified from B\'ar\'any \cite{barany2008} Theorem 4.1]
For any set of vector $x_1,...,x_n$ in $\R^d$ such that $\|x_i\| \le 1$ there exists a set of signs $s_1,...,s_n$ such that $||\sum_i s_i x_i || \le \sqrt{d}$.
\end{theorem}
\begin{proof}
Consider the feasible region for $\sum_i \alpha_i x_i = 0$ and all $\alpha_i \in [-1,1]$.
This feasible region is not empty because it contains the origin. 
Consider an extreme point $\alpha^*_i$ and set $s_i = 1$ w.p.\ $(1+\alpha^*_i)/2$ and $s_i = -1$ w.p.\ $(1+\alpha^*_i)/2$.
$$
\E[\|\sum s_i x_i\|^2] =  \E[\|\sum (s_i - \alpha^*_i) x_i\|^2] = \sum \E[(s_i - \alpha^*_i)^2] \|x_i\|^2= \sum (1-(\alpha_i^*)^2) \le d
$$
The last inequality holds because the values $\alpha^*_i$ are non-integer in at most $d$ places. 
Assuming otherwise would entail more than $d$ linearly independent vectors in $\R^d$ due to $\alpha^*$ being an extreme point.
\end{proof}

\section{Attempt at Cleaner Definitions}
Let $\mathcal F$ be a family functions from a domain $\X$ to $\R$.
We define the discrepancy of $\F$. Let $f \subset{\F}$ to a subset of $\F$.
$$\mathcal D_m(\F) := \sup_{x\in\X} \min_{s\in \{0,1\}^m} \sup_{f\subset \F, |f|=m}  \sum_{i}s_i f_i (x)$$  \el{need to double check the order here...}

We claim that a very common task is to approximate the function $F = \sum_i f_i$ using a concise representation.  
Specifically, using coresets we can approximate $F$ with $\tilde F = \sum_i w_i f_i$ where $w$ is non-zero in at most $k$ places.

In this paper we show that if a family of functions exhibits low discrepancy then finding a good coreset is always possible.
The size of the coresets we find with the proposed algorithm is always smaller than that achievable by uniform sampling.
Moreover, we show that this framework includes a wide variety of problems including quantile approximation, approximate counting, machine learning classification and regressing, kernel density estimation and many more. 
Finally, we show that coresets can always be created in a streaming fashion using fully mergable sketches with only a $\operatorname{poly}\log\log$ factor loss in space.

In what follows, we subscript the function $f$ with its parametrization.
\begin{itemize}
\item In approximate counting we have $f_a(x) = 1$ if $a=x$ and zero else. Here both $a$ and $x$ belong to some finite domain.
\item In quantile approximation we have $f_a(x) = 1$ if $a<x$ and zero else. Here, $a$ and $x$ belong to a set which exhibits total ordering.
\item In matrix column subset selection we have $f_z(a,q) = \langle z,x \rangle ^2$ where $a,x \R^d$. For simplicity, we assume $\|a\| = \|x\| = 1$.
\item In density estimation $f_a(x) = \exp(- \|a-x\|_2^2/\sigma^2)$ or any other kernel (see \cite{})
\item In logistic regression $f_a(x) = \operatorname{loss}(a,x) = \log(1 + \exp(-a^T x))$. 
Here $x$ is the parameters of the linear model and $a$ is a training example vector multiplied by its label ($1$ or $-1$).
\item Linear Classification $f_a(x) = 1$ if $a^Tx > 0$ and $0$ else. Here, like in Logistic Regression, $a$ is the example point multiplied by its label and $x$ is the weights of the linear classifier.
\item Fractional Satisfiability $f(x,q)$ if $q$ satisfies $f$ and $0$ otherwise. Here $x$ in a clause in conjunctive normal form and $q$ is a boolean value assignment to the variables. For example $f(x,q) = 1$ iff $(q_1 \vee q_5  \vee \neg q_8)$ is true.
\item Quadratic forms of the graph Laplacian are $f(x, q) = q^T L_x q = \sum_{i,j\in E} (q_i - q_i)^2$ where $x$ is an edge and $L_x$ is the graph laplacian corresponding to the graph with a single edge $x$ and $q$ is a test vector.
\item Probably also good for matrix approximation when we get entry updates. 
\end{itemize}

@@

\begin{definition}
Let $\F:\mathcal Q\rightarrow \R$ be a family of functions, $f \subset \F$, and $\sigma \in \{-1,1\}^m$ uniformly. 
The Rademacher complexity of $\F$ is defined to be
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f}  \sum_{i=1}^{m}\sigma_i f_i
\]
\end{definition}
\cite{Bartlett:2003:RGC:944919.944944}
\begin{definition}
Let $\F:\mathcal Q\rightarrow \R$ be a family of functions, $f \subset \F$, and $\sigma \in \{-1,1\}^m$.
The Komlos complexity of $\F$ is defined as
\[
K_m(\F) =  \frac{1}{m}\min_{\sigma} \sup_{f}  \sum_{i=1}^{m}\sigma_i f_i
\]
\end{definition}
We claim that while $R_n$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution) $K_n$ is the measure more suitable for
active learning and creating streaming coresets. 
The name Komlos Complexity due to the Komlos Conjecture which claims the minimal value of $\min_\sigma \max_f \|\sum_i \sigma_i f_i\|_\infty = O(1)$ for unit vectors $f$. The correctness of the Komlos Conjecture is still an important open problem in the study of discrepancy theory.

It is clear that $K_m(\F) < R_m(\F)$ for every family $\F$.

\newcommand{\ccc}[1]{\parbox}
\begin{table}[htp]
\caption{default}
\begin{center}
\begin{tabular}{|c|c|c|c|} \hline
Problem				& $f$ 											& $R_m$ 			& $K_m$ \\ \hline
Counting 				& $f(x, q) = 1$ if $x=q$ and zero else 					& $1/\sqrt{m}$		& $1/m$ \\ \hline
Quantiles				& $f(x, q) = 1$ if $x\le q$ and zero else  					& $1/\sqrt{m}$		& $1/m$ \\ \hline
Column Slection		& $f(x, q) = \langle x,q \rangle ^2$ where  $q,x\in \R^d$		& $1/\sqrt{m}$		& $\sqrt{d}/m$ \\ \hline
Kernel Density 			& $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ 					& $1/\sqrt{m}$ 		& $\sqrt{d}/m$ \\ \hline
Logistic Regression     	& $f(x, q) = \log(1 + \exp(q^T x))$ 						& $1/\sqrt{m}$ 		& $\sqrt{d}/m$ \\ \hline
Hinge Regression     		& $f(x, q) = [1-q^T x]_+$								& $1/\sqrt{m}$ 		& $\sqrt{d}/m$ \\ \hline
Linear Classification		& $f(x,q) = 1$ if $x^Tq > 0$ and $0$ else					& ?				& ? \\ \hline
Graph Laplacian 		& $f(x, q) = w_{ij}(q_i - q_i)^2$ for $x = (w_{ij},i,j)$, $q \in \R^n$ & ?				& ? \\ \hline
Matrix Approximation  	& $f(x, q) = w_{ij}u_iv_j$ for $x = (a_{ij},i,j)$  and $q = (u,v)$ 	& ?				& ? \\ \hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

>>>>

\section{Appendix 1}
For completeness we recap a result from \cite{barany2008}.
\begin{theorem}[Simplified from B\'ar\'any \cite{barany2008} Theorem 4.1]
For any set of vector $x_1,...,x_n$ in $\R^d$ such that $\|x_i\| \le 1$ there exists a set of signs $s_1,...,s_n$ such that $||\sum_i s_i x_i || \le \sqrt{d}$.
\end{theorem}
\begin{proof}
Consider the feasible region for $\sum_i \alpha_i x_i = 0$ and all $\alpha_i \in [-1,1]$.
This feasible region is not empty because it contains the origin. 
Consider an extreme point $\alpha^*_i$ and set $s_i = 1$ w.p.\ $(1+\alpha^*_i)/2$ and $s_i = -1$ w.p.\ $(1-\alpha^*_i)/2$.
$$
\E[\|\sum s_i x_i\|^2] =  \E[\|\sum (s_i - \alpha^*_i) x_i\|^2] = \sum \E[(s_i - \alpha^*_i)^2] \|x_i\|^2= \sum (1-(\alpha_i^*)^2) \le d
$$
The first equality is due to $\sum_i \alpha_i x_i = 0$. The second used $\E[s_i] = \alpha^*_i $.
The last inequality holds because the values $\alpha^*_i$ are non-integer in at most $d$ places. 
Assuming otherwise would entail more than $d$ linearly independent vectors in $\R^d$ due to $\alpha^*$ being an extreme point.
\end{proof}

%\section{Attempt at Cleaner Definitions}
%Let $\mathcal F$ be a family functions from a domain $\X$ to $\R$.
%We define the discrepancy of $\F$. Let $f \subset{\F}$ to a subset of $\F$.
%$$\mathcal D_m(\F) := \sup_{x\in\X} \min_{s\in \{0,1\}^m} \sup_{f\subset \F, |f|=m}  \sum_{i}s_i f_i (x)$$  \el{need to double check the order here...}
%
%We claim that a very common task is to approximate the function $F = \sum_i f_i$ using a concise representation.  
%Specifically, using coresets we can approximate $F$ with $\tilde F = \sum_i w_i f_i$ where $w$ is non-zero in at most $k$ places.
%
%In this paper we show that if a family of functions exhibits low discrepancy then finding a good coreset is always possible.
%The size of the coresets we find with the proposed algorithm is always smaller than that achievable by uniform sampling.
%Moreover, we show that this framework includes a wide variety of problems including quantile approximation, approximate counting, machine learning classification and regressing, kernel density estimation and many more. 
%Finally, we show that coresets can always be created in a streaming fashion using fully mergable sketches with only a $\operatorname{poly}\log\log$ factor loss in space.
%
%
%In what follows, we subscript the function $f$ with its parametrization.
%\begin{itemize}
%\item In approximate counting we have $f_a(x) = 1$ if $a=x$ and zero else. Here both $a$ and $x$ belong to some finite domain.
%\item In quantile approximation we have $f_a(x) = 1$ if $a<x$ and zero else. Here, $a$ and $x$ belong to a set which exhibits total ordering.
%\item In matrix column subset selection we have $f_z(a,q) = \langle z,x \rangle ^2$ where $a,x \R^d$. For simplicity, we assume $\|a\| = \|x\| = 1$.
%\item In density estimation $f_a(x) = \exp(- \|a-x\|_2^2/\sigma^2)$ or any other kernel (see \cite{})
%\item In logistic regression $f_a(x) = \operatorname{loss}(a,x) = \log(1 + \exp(-a^T x))$. 
%Here $x$ is the parameters of the linear model and $a$ is a training example vector multiplied by its label ($1$ or $-1$).
%\item Linear Classification $f_a(x) = 1$ if $a^Tx > 0$ and $0$ else. Here, like in Logistic Regression, $a$ is the example point multiplied by its label and $x$ is the weights of the linear classifier.
%\item Fractional Satisfiability $f(x,q)$ if $q$ satisfies $f$ and $0$ otherwise. Here $x$ in a clause in conjunctive normal form and $q$ is a boolean value assignment to the variables. For example $f(x,q) = 1$ iff $(q_1 \vee q_5  \vee \neg q_8)$ is true.
%\item Quadratic forms of the graph Laplacian are $f(x, q) = q^T L_x q = \sum_{i,j\in E} (q_i - q_i)^2$ where $x$ is an edge and $L_x$ is the graph laplacian corresponding to the graph with a single edge $x$ and $q$ is a test vector.
%\item Probably also good for matrix approximation when we get entry updates. 
%\end{itemize}
%

@@

\begin{definition}
Let $\F:\mathcal Q\rightarrow \R$ be a family of functions, $f \subset \F$, and $\sigma \in \{-1,1\}^m$ uniformly. 
The Rademacher complexity of $\F$ is defined to be
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f}  \sum_{i=1}^{m}\sigma_i f_i
\]
\end{definition}
\cite{Bartlett:2003:RGC:944919.944944}
\begin{definition}
Let $\F:\mathcal Q\rightarrow \R$ be a family of functions, $f \subset \F$, and $\sigma \in \{-1,1\}^m$.
The Komlos complexity of $\F$ is defined as
\[
K_m(\F) =  \frac{1}{m}\min_{\sigma} \sup_{f}  \sum_{i=1}^{m}\sigma_i f_i
\]
\end{definition}
We claim that while $R_n$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution) $K_n$ is the measure more suitable for
active learning, for creating coresets, and for creating mergeable sketches.
The name Komlos Complexity due to the Komlos Conjecture. It claims that $\min_\sigma \max_f \|\sum_i \sigma_i f_i\|_\infty = O(1)$ for unit vectors $f$. The correctness of the Komlos Conjecture is still a fundamental open problem in the study of discrepancy theory. 
The definition is useful, however regardless of the correctness of the conjecture.  
It is obvious that  $K_m(\F) \le R_m(\F)$ for any family $\F$. We claim, however, that $K_m(\F) = o(R_m(\F))$ for a wide range of interesting problems in machine learning.

\newcommand{\ccc}[1]{\parbox}
\begin{table}[htp]
\caption{default}
\begin{center}
\begin{tabular}{|c|c|c|c|} \hline
Problem				& $f$ 											& $R_m$ 			& $K_m$ \\ \hline
Counting 				& $f(x, q) = 1$ if $x=q$ and zero else 					& $1/\sqrt{m}$		& $1/m$ \\ \hline
Quantiles				& $f(x, q) = 1$ if $x\le q$ and zero else  					& $1/\sqrt{m}$		& $1/m$ \\ \hline
Column Slection		& $f(x, q) = \langle x,q \rangle ^2$ where  $q,x\in \R^d$		& $1/\sqrt{m}$		& $\sqrt{d}/m$ \\ \hline
Kernel Density 			& $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ 					& ? 		& $\sqrt{m}$ \\ \hline
Logistic Regression     	& $f(x, q) = \log(1 + \exp(q^T x))$ 						& $1/\sqrt{m}$ 		& $\sqrt{d}/m$ \\ \hline
Hinge Regression     		& $f(x, q) = [1-q^T x]_+$								& $1/\sqrt{m}$ 		& $\sqrt{d}/m$ \\ \hline
Linear Classification		& $f(x,q) = 1$ if $x^Tq > 0$ and $0$ else					& ?				& ? \\ \hline
Graph Laplacian 		& $f(x, q) = w_{ij}(q_i - q_i)^2$ for $x = (w_{ij},i,j)$, $q \in \R^n$ & ?				& ? \\ \hline
Matrix Approximation  	& $f(x, q) = w_{ij}u_iv_j$ for $x = (a_{ij},i,j)$  and $q = (u,v)$ 	& ?				& ? \\ \hline
\end{tabular}
\end{center}
\label{default}
\end{table}%
